High-order cognition is supported by complex but compressible brain activity patterns

### 2023-02-23 paper/main.tex
checking in a draft of the introduction

% % compression and task demands -- MackEtal20, ZhouEtal19
% % timecorr -- higher order network patterns during higher order cognition implies more complexity
% % cognitive richness and scrambling-- various Hasson lab papers

% We're interested in the complexity of brain patterns that underly different
% types of thoughts. To explore this question space, we will take brain patterns
% recorded under different experimental conditions used in Aim 2, and project
% them into lower dimensional spaces using principle components analysis. We can
% then ask how well those low-dimensional embeddings of the data retain
% cognitively relevant information like when in a story someone is listening to.

% This work has been inspired, in part, by~\cite{MackEtal20}. In this paper, they
% investigated the role of the prefrontal cortex in filtering out irrelevant
% content. Specifically, they looked at if the vmPFC performs data reduction on
% incoming information through compression. This was motivated, in part, by
% orbital frontal cortex (OFC) compression in rats~\citep{ZhouEtal19}. They
% studied this using a learning paradigm in which participants had to classify
% insects based on different numbers of feature dimensions. The idea was that
% participants in some learning blocks, participants could identify the insects
% based on one feature (low complexity) or several features (high complexity),
% but importantly the stimuli remained the same across all learning problems.
% They found that complexity and compression had an inverse relationship; the
% lower complexity of a conceptual space, the higher the degree of compression.
% Building on this idea, we wonder if varying degrees of compression is performed
% throughout the brain. We also want to test this idea, but using varying levels
% of engagement listening to a naturalistic stimuli.

% To understand the degree of compression throughout the brain during cognition,
% we will use the same fMRI data from Aim 2, collected while participants
% listened to a story in different scrambling conditions. We will measure the
% degree that multivoxel activation patterns are compressed during story
% listening using principle components analysis (PCA) a method for low-rank
% approximation of multidimensional data~\citep{EckaYoun36}. We will explore this
% using decoding accuracy as a function of the number of components, or
% dimensions, in the low-dimensional space under different cognitive conditions.

% You can imagine two reasonable predictions of how cognition is reflected in
% brain patterns. The first is as our thoughts become more complex, they are
% supported by more complex brain patterns, and require more components to
% decode. The second is that when thoughts are deeper and more complicated, the
% units of neural activity would carry more information, and would require
% therefore fewer components to decode.

% This idea can be explored in this visual analogy (Fig.~\ref{fig: pie_example})
% for neural compression. Here there are two images of pies, the top pie is more
% complex than the bottom. On the left we're illustrating that it takes fewer
% components to reach the same 95 percent variance explained in the less complex
% pie, which corresponds to higher compression. However, on the right with very
% few components similar variance is explaining both pies.

% We investigated the dimensionality of neural patterns by training
% classifiers using more and more principle components. Or, in other
% words, we used less and less compression to decode.
% We applied the approach to
% a neuroimaging dataset comprising data collected as participants
% listened to a story varying in cognitive richness~\citep{SimoEtal16}.

>>>>

The above prior studies are consistent with two potential descriptions of how
cognitive processes are reflects in brain activity patterns. One possibility is
that the information rate of brain activity increases during more complex or
higher-level cognitive processing. If so, then the ability to reliably decode
cognitive states from brain activity patterns should improve with task
complexity or with the level (or ``depth'') of cognitive processing. A second
possibility is that the compressibility of brain activity patterns increases
during more complex or higher-level cognitive processing. If so, then
individual features of brain recordings, or compressed representations of brain
recordings, should carry more information during complex or high-level (versus
simple or low-level) cognitive tasks.

We used a previously collected neuroimaging dataset to estimate the extent to
which each of these two possibilities might hold. The dataset we examined
comprised functional magnetic resonance imaging (fMRI) data collected as
participants listened to an audio recording of a 10-minute story, temporally
scrambled recordings of the story, or underwent a resting state
scan~\citep{SimoEtal16}. Each of these experimental conditions evokes different
depths of cognitive processing~\citep{SimoEtal16,LernEtal11,
HassEtal08,OwenEtal21}. We used across-participant classifiers to decode
listening times in each condition, as a proxy for how ``informative'' the
task-specific activity patterns were~\citep{SimoChan20}. We also use principle
components analysis to generate lower-dimensional representations of the
activity patterns. We then repeated the classification analyses after
preserving different numbers of components and examined how classification
accuracy changed across the different experimental conditions.
